---
title: "Data Analysis"
author: "Gina Yi"
date: '2022-11-20'
output: html_document
---
Libraries Used:
```{r}
library("ggplot2")
library("tidyverse")
library("car")
```

Data Cleaning:
Responses to remove:
"ANXFREQ_A": 7-9
"DEPFREQ_A": 7-9
"PHSTAT_A': 7-9 
"HISPALLP_A":97-99 (None of these values exist in the data set)
"MARSTAT_A": 9
"ORIENT_A": 7-9
```{r}
datUnclean <- read.csv("~/Desktop/4960 Project/datUnclean.cvs")
dat <- datUnclean
colnames(dat) <- c("x","anxiety","depression","physicalHealth","raceEthnicity", "marital", "sexuality")
dat$anxiety <- ifelse(dat$anxiety >=7, NA, dat$anxiety)
dat$depression <- ifelse(dat$depression >=7, NA, dat$depression)
dat$physicalHealth <- ifelse(dat$physicalHealth >= 7, NA, dat$physicalHealth)
dat$marital <-  ifelse(dat$marital == 9, NA, dat$marital)
dat$sexuality <- ifelse(dat$sexuality >= 7, NA, dat$sexuality)
```

Relevel and Reverse Coding 
```{r}
dat$anxietyRev<-car::recode(dat$anxiety, "1=5;2=4;3=3;4=2;5=1")
dat$depressionRev<-car::recode(dat$depression, "1=5;2=4;3=3;4=2;5=1")
dat$physicalHealthRev<-car::recode(dat$physicalHealth, "1=5;2=4;3=3;4=2;5=1")
dat$raceEthnicity<-as.factor(dat$raceEthnicity)
levels(dat$raceEthnicity)<-c("Hispanic", "Non-Hispanic White only", "Non-Hispanic Black/African American only", "Non-Hispanic Asian only", "Non-Hispanic AIAN only","AIAN and any other group","Other single and multiple races")
dat$marital<-as.factor(dat$marital)
levels(dat$marital)<-c("Married","Married","Married","Widowed","Divorced","Separated","Never married","Living with partner")
# dat$physicalHealthRev<-as.factor(dat$physicalHealthRev)
# levels(dat$physicalHealthRev)<-c("Poor","Fair","Good","Very Good","Excellent")

```

```{r}
"Currently Unused"
#dat$physicalHealthRev <- as.ordered(dat$physicalHealth)
#levels(dat$physicalHealthRev) <- rev(levels(dat$physicalHealthRev))

#dat$depressionRev <- as.ordered(dat$depression)
#levels(dat$depressionRev) <- rev(levels(dat$depressionRev))

#dat$anxietyRev<- as.ordered(dat$anxiety)
#levels(dat$anxietyRev)<- rev(levels(dat$anxietyRev))

```


Physical Health Hypotheses:
Individuals with poorer physical health statuses will experience depression more frequently.
```{r}
darPhysStatDep <- dat %>% select("physicalHealthRev", "depressionRev")
darPhysStatDep <- na.omit(darPhysStatDep)

darPhysStatAnx <- dat %>% select("physicalHealthRev", "anxietyRev")
darPhysStatAnx <- na.omit(darPhysStatAnx)
```

```{r}
cor.test(dat$depressionRev, dat$physicalHealthRev, method = "spearman", alterantive = "greater")
# ggplot(dat, aes(depressionRev,color=as.factor(physicalHealthRev),fill=as.factor(physicalHealthRev))) + geom_bar(alpha=.5,position = position_dodge()) + xlab("Depression Score") + ylab("Responses") + labs(color = "Physical Health", fill = "Physical Health")

ggplot(darPhysStatDep, aes(depressionRev,color=as.factor(physicalHealthRev),fill=as.factor(physicalHealthRev))) + geom_bar(alpha=.5,position = position_dodge()) + xlab("Depression Scores") + ylab("Responses") + labs(color = "Physical Health", fill = "Physical Health") + scale_fill_discrete(labels = c("Poor", "Fair", "Good", "Very Good", "Excellent")) + scale_color_discrete(labels = c("Poor", "Fair", "Good", "Very Good", "Excellent"))
```

Individuals with poorer physical health statuses will experience anxiety more frequently.
```{r}
cor.test(dat$physicalHealthRev, dat$anxiety, method = "spearman", alternative = "greater")
# ggplot(dat, aes(anxietyRev,color=as.factor(physicalHealthRev),fill=as.factor(physicalHealthRev))) + geom_bar(alpha=.5,position = position_dodge())

ggplot(darPhysStatAnx, aes(anxietyRev,color=as.factor(physicalHealthRev),fill=as.factor(physicalHealthRev))) + geom_bar(alpha=.5,position = position_dodge()) + xlab("Anxiety Scores") + ylab("Responses") + labs(color = "Physical Health", fill = "Physical Health") + scale_fill_discrete(labels = c("Poor", "Fair", "Good", "Very Good", "Excellent")) + scale_color_discrete(labels = c("Poor", "Fair", "Good", "Very Good", "Excellent"))
```


Race/Ethnicity Hypotheses:
Individuals of different races will experience different frequencies of depression.
```{r}
kruskal.test(depressionRev ~ raceEthnicity, data= dat)
ggplot(dat, aes(depressionRev,color=raceEthnicity,fill=raceEthnicity)) + geom_bar(alpha=.5, position = position_dodge())+ylab("Responses")+xlab("Depression Score")+ labs(color= "Racial/Ethnic Groups", fill="Racial/Ethnic Groups")
```

Individuals of different races will experience different frequencies of anxiety.
```{r}
kruskal.test(anxietyRev ~ raceEthnicity, data= dat)
ggplot(dat, aes(anxietyRev,color=raceEthnicity,fill=raceEthnicity)) + geom_bar(alpha=.5, position = position_dodge())+ylab("Responses")+xlab("Anxiety Score")+ labs(color= "Racial/Ethnic Groups", fill="Racial/Ethnic Groups")
```


Marital status Hypotheses:
Individuals who are married will experience lower levels of of depression
```{r}
datMarStatAnx<-dat %>% select("marital","anxietyRev")
datMarStatDep<- dat %>% select("marital","depressionRev")
datMarStatAnx<-na.omit(datMarStatAnx)
datMarStatDep<- na.omit(datMarStatDep)
```


```{r}
kruskal.test(depressionRev ~ marital, data= datMarStatDep)
ggplot(datMarStatDep, aes(depressionRev,color=marital,fill=marital)) + geom_bar(alpha=.5,position = position_dodge())+xlab("Depression Score")+ ylab("Responses")+labs(color="Marital Status",fill="Marital Status")
```


Individuals who are married will experience lower levels of of anxiety
```{r}
kruskal.test(anxietyRev ~ marital, data= datMarStatAnx)
ggplot(datMarStatAnx, aes(anxietyRev,color=marital,fill=marital)) + geom_bar(alpha=.5,position = position_dodge())+xlab("Anxiety Score")+ ylab("Responses")+labs(color="Marital Status",fill="Marital Status")
```

