---
title: "Data Analysis"
author: "Gina Yi"
date: '2022-11-20'
output: html_document
---
Libraries Used:
```{r}
library("tidyverse")
```

Data Cleaning:
Responses to remove:
"ANXFREQ_A": 7-9
"DEPFREQ_A": 7-9
"PHSTAT_A': 7-9 
"HISPALLP_A":97-99 (None of these values exist in the data set)
"MARSTAT_A": 9
"ORIENT_A": 7-9
```{r}
datUnclean <- read.csv("~/Desktop/PSY4960-Fall2022/PSY4960Project/datUnclean.csv")
dat <- datUnclean
colnames(dat) <- c("x","anxiety","depression","physicalHealth","raceEthnicity", "marital", "sexuality")
dat$anxiety <- ifelse(dat$anxiety >=7, NA, dat$anxiety)
dat$depression <- ifelse(dat$depression >=7, NA, dat$depression)
dat$physicalHealth <- ifelse(dat$physicalHealth >= 7, NA, dat$physicalHealth)
dat$marital <-  ifelse(dat$marital == 9, NA, dat$marital)
dat$sexuality <- ifelse(dat$sexuality >= 7, NA, dat$sexuality)
```
Relevel and Reverse Coding 
```{r}
physicalHealth1 <- as.ordered(dat$physicalHealth)
dat$physicalHealthRev <- as.factor(dat$physicalHealth)
levels(dat$physicalHealthRev) <- rev(levels(dat$physicalHealthRev))
dat$depressionRev <- as.factor(dat$depression)
levels(dat$depressionRev) <- rev(levels(dat$depressionRev))
```


Physical Health Hypotheses:
Individuals with poorer physical health statuses will experience depression more frequently.
```{r}
cor.test(dat$depression, dat$physicalHealth, method = "spearman", alterantive = "greater")
ggplot(data = dat, aes(physicalHealth, depression)) + geom_point()
```

Individuals with poorer physical health statuses will experience anxiety more frequently.
```{r}
cor.test(dat$physicalHealth, dat$anxiety, method = "spearman", alternative = "greater")
```

Race/Ethnicity Hypotheses:
Individuals of different races will experience different frequencies of depression.
```{r}

```

Individuals of different races will experience different frequencies of anxiety.
```{r}

```

Marital status Hypotheses:
Individuals who are married will experience lower levels of of depression
```{r}

```

Individuals who are married will experience lower levels of of anxiety
```{r}

```

